\input{header.tex}

\usepackage{clrscode}
\usepackage{enumerate}
\newcommand{\hmwkTitle}{Assignment\ \#5} % Assignment title
\newcommand{\hmwkDueDate}{April\ 4,\ 2016} % Due date
\newcommand{\hmwkClass}{Algorithms} % Course/class
\newcommand{\hmwkAuthorName}{Zhaoyang Li (2014013432)} % Your name
%----------------------------------------------------------------------------------------

\begin{document}

\maketitle

%----------------------------------------------------------------------------------------
%	TABLE OF CONTENTS
%----------------------------------------------------------------------------------------

\setcounter{tocdepth}{1} % Uncomment this line if you don't want subsections listed in the ToC

\newpage
\tableofcontents
\newpage

%----------------------------------------------------------------------------------------
%	PROBLEM 1
%----------------------------------------------------------------------------------------


\begin{homeworkProblem}

Read CLRS chapters 12, 13 and 18.

\problemAnswer{
}

\end{homeworkProblem}


%----------------------------------------------------------------------------------------
%	PROBLEM 2
%----------------------------------------------------------------------------------------


\begin{homeworkProblem}

CLRS Exercises 17.4-3.


\problemAnswer{

Let $c_i$ be actual cost of deletion operation $i$, and $\hat{c_i}$ be the corresponding amortized one. Let $n_{i-1}, s_{i-1}, n_i, s_i$ be number of elements in the table and size of the table, before and after the deletion, respectively.

When a deletion happens without resizing,

$$
s_i = s_{i-1},
n_i = n_{i-1} - 1,
n_i p\geq \frac13 s_{i-1},
$$

Therefore

\begin{equation}
\begin{aligned}
\hat{c_i}&=c_i+\Phi_i-\Phi_{i-1} \\
&=1+|2 n_i - s_i| - |2 n_{i-1} - s_{i-1}| \\
&\leq1+|\left(2 \left(n_{i-1}-1\right) - s_{i-1}\right) - \left(2 n_{i-1} - s_{i-1}\right)| \\
&=1+2\\
&=3.
\end{aligned}
\end{equation}

When a resizing happens, we have

$$
s_i = \lfloor\frac23 s_{i-1}\rfloor,
n_i = n_{i-1} - 1,
n_{i-1} \geq\frac13 s_{i-1},
n_{i} < \frac13 s_{i-1}
$$

Therefore
\begin{equation}
\begin{aligned}
\hat{c_i}&=c_i+\Phi_i-\Phi_{i-1} \\
&=\left(n_i+1\right) +|2 n_i - s_i| - |2 n_{i-1} - s_{i-1}| \\
&=\left(n_i+1\right) + |s_i - 2 n_i| -\left(-2 n_{i-1} + s_{i-1}\right)\\
&\leq\left(n_i+1\right) + 2 - n_i \\
&=3.
\end{aligned}
\end{equation}

Therefore

$$\hat{c_i}=O(1)$$

QED.
}

\end{homeworkProblem}

%----------------------------------------------------------------------------------------
%	PROBLEM 3
%----------------------------------------------------------------------------------------


\begin{homeworkProblem}

CLRS Problems 17-2.
Making binary search dynamic.

a.

\problemAnswer{
To search, perform standard binary search on each of the $k$ sorted arrays.

Worst-case running time
\begin{equation}
\begin{aligned}
T(n)
& \leq\sum_{i=0}^{k-1}O(\lg{2^i}) \\
& =\sum_{i=0}^{k-1}O(i) \\
& =O\left(\sum_{i=0}^{k-1}i\right) \\
& =O\left(k^2\right) \\
& =O\left(\lceil\lg(n+1)\rceil^2\right)\\
& =O\left(\lg^2n\right)
\end{aligned}
\end{equation}
}

b.

\problemAnswer{
To insert an element $e$, 
\begin{quote}
\begin{enumerate}[Step 1.]
\item Find the max $j$ such that $\forall i\ s.t\ 0\leq i<j, n_i = 1$ and $n_j=0$. If such $j$ does not exist, we say $j=0$.
\item Let $A_j=\bigcup_{i=0}^{j-1}A_{i}$. 
\item Insert $e$ into $A_j$.
\item For all $i$ such that $0\leq i < j$, let $A_{i}=\langle\rangle, n_i=0$.
\item Let $n_j = 1, n = n+1$. Done.
\end{enumerate}
\end{quote}

Timing:\\ \\ 
Steps 1, 6 take $O(1)$ time. Step 3 takes time linear to length of $A_j$, which is $O(2^j)$. Step 4 takes $O(j)$. Step 2 involves merging sorted arrays and can be done in $O(2^j)$ time. The total is $O(2^j)$.
\\ \\ 
In the worst case, $j=k-1$, meaning $n=2^k-1$. Running time $O(2^j)=O(2^k)=O(2^{\lg n}) = O(n)$.
\\ \\ 
During $m$ insertion operations, noticing the nature of binary increment, in $\frac m2$ cases we have $j=0$, in $\frac m4$ for $j=1$, $\frac m8$ for $j=2$... In a word, we know the distribution of $j$.

$$T_m
\leq\sum_{i=0}^{k-1}\frac{m}{2^i} O(\lg{2^{i+1}}) 
\leq\sum_{i=0}^{k-1}m O(1)
=mO(k)
=mO(\lg n)$$

Applying aggregate analysis, amortized running time for each operation 



$$T = \frac{m O(\lg n)}{m} = O(\lg n)$$
}

c.

\problemAnswer{
To perform a deletion of element $e$, we first perform a search to find $i$ s.t. $e\in A_i$. Remove $e$ from $A_i$. 

Also, we find the minimal $j$ s.t. $A_j\neq \langle\rangle$. Take an element $a\in A_j$ randomly and insert it into $A_i$. Now forget about $A_i$ and look at $A_j$, whose length is $2^j-1$. We divide $A_j$ into arrays $A_0, A_1, ..., A_{j-1}$ who have a total capacity of $2^j-1$ and are by now empty. Done.
}
\end{homeworkProblem}


%----------------------------------------------------------------------------------------
%	PROBLEM 4
%----------------------------------------------------------------------------------------


\begin{homeworkProblem}

CLRS Problems 19-3. More Fibonacci-heap operations.

a.

\problemAnswer{
If $k < x.key$, we call \proc{Fib-Heap-Decrease-Key}$(H, x, k)$, amortized cost of which is O(1).

If $k = x.key$, we do nothing and return, with $O(1)$ time consumption.


If $k > x.key$, call \proc{Cut}$(H, y, y.p)$ for each child of $x$ such that $y.p = x$. Then let $x.key = k$, call \proc{Cut}$(H, x, x.p)$. Done. Since $\forall z \in H, z.degree=O(\lg n)$, we have the amortized cost $O(\lg n)$.
}

b.

\problemAnswer{
Delete leaves only. We may want to keep track of all the leaves, so that deletion of a single leaf takes $O(1)$ time.

Let $\Phi(D_i)=t(H) + m(H)$.

$\hat{c_i} = c_i +  \Phi(D_i) - \Phi(D_{i-1})
= q O(1) - q
= O(q)
$.
}
\end{homeworkProblem}



%----------------------------------------------------------------------------------------

\end{document}
